R Notebook

### 2022-02-16 docs/3. SNA - Assignments/sna_Week3/Week3_Notebook.Rmd
whelp, there's data in the statnet and igraph object, but sure if it's meaningful though

@@

```{r EDGELIST: Network Objects}

@@

    View(network_nodes)
######
      ######**igraph** 
######
        #igraph object 
network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx), 
    directed=F)
                ## create new edge attribute
  E(network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(network_ig)$ingType<-smugCovIdx$ingType
  #create vertex attributes in igraph
 V(network_ig)[match(V(smug_matrix)$dnkName,network_nodes$dnkName)]$dnkName<-
   network_nodes$dnkName

  V(network.ig)[match(V(network_edgelist)$name,network_nodes$VERTEX_NAME)]$dnkName <- network_nodes$dnkName
      summary(edge_attr(network_ig))

head(network_ig)
summary(network_ig)
      ######

@@

summary(network_statnet)

                                 #     access vertex attributes
  view(network_statnet%v% "vertex.names")
            # list the names of vertex attributes
  network::list.vertex.attributes(network_statnet)
            #  Generate a list the names of edge attributes
  network::list.edge.attributes(network_statnet)
            #   summarize numeric network attribute
  summary(network_statnet%v%"dnkName")

@@

## Smuggler's Cove Index- NETWORK: EDGELIST: Features & Descriptions
```{r Network: Identify critical features, echo=FALSE}

#**Statnet EdgeList**   smugEL_sn
#####

```{r STATNET: Network Description}

@@

      # list the structure of smugCovIdx
str(smugEL_sn )

      ## Network Attributes
 print(smugEL_sn)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names: 
```

@@

## Network Size
      # Count Vertices
    vcount(smugEL_ig)

@@

**Vertex and Edge Attributes**

```{r statnet: Attribute Data: Directed: Adjacency Network, echo=TRUE}

# class of an object
    class(smugEL_sn)

#     access vertex attributes
  view(smugEL_sn %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(smugEL_sn)

#     access edge attribute
#  head(smugEL_sn%e% "weight")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(smugEL_sn)

#   summarize numeric network attribute
  summary(smugEL_sn  %v% "name")

```{r igraph: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attribute 
  V(smugEL_ig)$name
#     Generate a list the names of vertex attributes
  igraph::vertex_attr_names(smugEL_ig)

#     access edge attribute
  E(smugEL_ig)$weight

@@

#   Week 2: Scripts
##    Network: Describe Network Data
##    *Identify and describe content of nodes and
    links,*

Vertex attribute names: vertex.names

3.  *Identify whether attribute data are present.*

```{r Summary,echo=TRUE}

  summary(smugCovIdx,
          print.adj = TRUE
          )
```

#       Explore the dataset using commands from week 2 tutorial.

5.  *geodesic and path distances,*

@@

```{r iGraph:Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(smugEL_ig,"node 1","node 2")

#     Calculate distance between two nodes using unweighted edges
    distances(smugEL_ig,"node 1", "node 5",weights=NA)

We can also find all of the shortest paths between
two famalies in this network. These list the
starting node and the ending node with all nodes
inbetween for each path. The option weights=NA
means that any available edge weights are ignored.

```{r iGraph:Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(smugEL_ig,"node 3","node 4", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

The concept of shortest path to describe the
overall network structure can be more useful

@@

6.  *triads or transitivity,*

## Dyad Census

Classifies all dyads in the network as: -
Reciprocal (mutual), or 
mut - Asymmetric (non-mutual), or 
asym, and - 
Absent, or null

@@

#     Classify all Triad in the network: Directed Network

  sna::triad.census(smugEL_sn, mode= )

@@

#     weighted network transitivity:
#   igraph::transivity(smugEL_ig)

@@

## Local Transivity or Clustering

`Local transitivity` (local clustering
coefficient), is a technical description of the
density of an ego network.

@@

#     Retrive a list of the vertices we are interested
    V(smugEL_ig)[c("node 1" ,
                "node 5", 
                "node 6",
                "node 2")]

#     check ego network transitivity
transitivity(smugEL_ig,
             type="local", 
             vids=V(smugEL_ig)
             [c("node 1",
                "node 5", 
                "node 6",
                "node 2")]) 
#     get global clustering cofficient
  transitivity(smugEL_ig, type="global")

@@

7.  *connectedness and.or component structure,
    etc.*

>>>>

#   Project Varriables

    Raw Data:
  indx_SmugglerCove_recip

    DataFrame:
  smugCovIdx
  network_nodes <- smugCovIdx << "dnkName"

    Matrix:
  smug_matrix
  network_edgelist <- smug_matrix

    Network Objects: EDGE LIST
  STATNET: smugEL_sn
  IGRAPH: smugEL_ig

@@

```{r EDGELIST: Network Objects: statnet & igraph, include=FALSE}

@@

        ### View(network_nodes)

######
      ######**igraph** 
######
        #igraph object 
network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[1:2]), 
    directed=F)
                ## create new edge attribute
  E(network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(network_ig)$ingType<-smugCovIdx$ingType
  #create vertex attributes in igraph
 V(network_ig)[match(V(network_ig)$dnkName,network_nodes$dnkName)]$dnkName<-
   network_nodes$dnkName

      summary(edge_attr(network_ig))

  # head(network_ig)
  # summary(network_ig)
      ######

@@

       # list the names: vertex attributes
  view(network_statnet%v% "vertex.names")

summary(network_statnet)
        #####

 ## Network Objects: EDGE LIST
              ##**STATNET:**
  smugEL_sn <- network_statnet

                ##**IGRAPH:** 
  smugEL_ig <- network_ig

@@

## Smuggler's Cove Index- NETWORK: EDGELIST: Features & Descriptions

```{r STATNET: Network Description, include=FALSE}

@@

## access vertex attributes
            # list the names:  vertex attributes
  network::list.vertex.attributes(smugEL_sn)

            # list the names:  edge attributes
  network::list.edge.attributes(smugEL_sn)
      #   summarize: numeric network attribute
  summary(smug_sn%v%"dnkName")

## Network Attributes
 print(smugEL_sn)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

##**NOTE** large quantity of output

  # list the structure of smugCovIdx
# str(smugEL_sn)
```

smugEL_sn

[1] "network"

Network attributes:
  vertices = 274 
  directed = TRUE 
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

@@

## Network Size

      # Count Vertices
    vcount(smugEL_ig)

@@

    # edge weights
    E(smugEL_ig)$weight

## access vertex attributes

    # Generate a list the names of vertex attributes
  V(smugEL_ig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(smugEL_ig)

@@

## Network Attributes
 summary(smugEL_ig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names: 
```
smugEL_ig

[1] "igraph"

Network attributes:
  vertices = 274 
  directed = FALSE
  hyper = FALSE 
  loops = FALSE 
  multiple = TRUE 
  bipartite = FALSE 
  total edges= 672 
    missing edges= 0 
    non-missing edges= 672

 Vertex attribute names: 
    dnkName vertex.names

 Edge attribute names: 
    dnkIng ingType

#   Week 2: Basic Network Structure: Smuggler's Cove Drink Index

@@

@@

```{r iGraph: Specific Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(smugEL_ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

```{r iGraph: Specific Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(smugEL_ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(smugEL_ig,"Bischeri", "Castellani",weights=NA)

## Dyad Census
Classifies all dyads in the network as: - 
Reciprocal (mutual), 
or mut - Asymmetric (non-mutual), 
or asym, 
and - Absent, or null

@@

@@

@@

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.

@@

#     Retrive a list of the vertices we are interested
    V(smugEL_ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(smugEL_ig,
             type="local", 
             vids=V(smugEL_ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(smugEL_ig, type="global")

@@

## Component Structure and Isolates

```{r Statnet: Isolates, echo=TRUE}

              ## Isolates:

 #List the isolates
 isolates(smugEL_sn)

 # Returns a list of the isolates in the graph or graph set
 isolates(smugEL_sn, diag=FALSE)

 # Is Ego an Isolate?
 is.isolate(smugEL_sn, ego, g=1, diag=FALSE)

 # Adds n isolates to the graph (or graphs)
 add.isolates(smugEL_sn, n, return.as.edgelist = FALSE)

```{r igraph: Component Structure, echo=TRUE}
                ## Component Structure:

    # Number of components
 igraph::components(smugEL_ig)$no

  # Size of each component
 igraph::components(smugEL_ig)$csize

## Network Density

 The density of a network is defined as the proportion of ties present out of all possible ties in a network.

 **Note:**

 the network density command ignores edge values at present.

 If loops are present or there are multiple edges between two nodes, then the results may not make sense.

```{r Statnet, echo=TRUE}

# get network density: with possible loops
 network.density(smugEL_sn)

 # get network density without loops: statnet
 gden(smugEL_sn, diag=FALSE)

 **Igraph** provides an incorrect density calculation as the default assumes  that loops are not included but does not remove them.

 Therefore, the default calculation includes the loops in the numerator but not the denominator of the density calculation.

```{r igraph: network density, echo=TRUE}
#get network density without loops: igraph
 network.density(smugEL_ig,loops=TRUE)

 # get network density
 network.density(smugEL_ig)
 # provides an incorrect density calculation as the default assumes that loops are not included but does not remove them

## Vertex Degree
 **Vertex degree** is a description of an individual node; each vertex in a network may have a different node degree.

 *Node Degree:*
   node degree is one measure of node centrality or popularity; 
 Node degree cis a count of the relationships that a node is involved in or the count of the edges that are incident to a vertex

 **use the network degree command to:**
   *  compute network summary statistics (such as average network degree);
 *  graph the network degree distribution; or
 *  identify particular nodes (such as the five nodes with the most contacts).

```{r Statnet: Vertex Degree, echo=TRUE}

      # statnet: Calculate average network degree

  sna::degree(smugEL_sn)
          # ignores loops or self-nominations

```{r igraph: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

 # includes loops
 igraph::degree(smugEL_ig)

## Network Degree Summary Statistics

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(network_nodes)
      # provides range, mean, median and standard deviation of a vector.

 # get summary statistics for node degree for the binary network degree

 summary(data.frame(
   name=smugEL_sn%v%"vertex.names",
   totdegree=sna::degree(smugEL_sn),
   indegree=sna::degree(smugEL_sn, cmode="indegree"),
   outdegree=sna::degree(smugEL_sn, cmode="outdegree")
 ))

#Visualizations

## igraph Visualizations
igraph provides three different ways for visualization.
```{r igraph: , echo=TRUE}

plot.igraph(smugEL_ig)

 # uses regular R graphics and can be used with any R device.

#  (Tk is quite resource hungry, so don't try this for very large graphs.)

 tkplot(smugEL_ig)

```{r Statnet: Visualization, echo=TRUE}

 # add in vertex labels and change the size and color of these labels
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75, 
       label.col="black"
 )

 # For directed graphs, you can turn off the arrows
 gplot(smugCovIdx, 
       displaylabels=TRUE, 
       label.cex=.75, 
       label.col="black", 
       usearrows = FALSE)

 # differentiate nodes based on their attributes
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.col=att$dnkName
 )

 legend("bottomleft",
        fill=0:4,
        legend=paste("DEPT",0:4),
        cex=0.75
 )

 # changed the shape of the nodes based on the level of their positions. 
 ## You change the shape of the nodes by specifying the number of sides you want the shape to have using the vertex.sides argument. 
 gplot(smugEL_sn, 
       displaylabels=TRUE,
       label.cex=.75,
       label.col="black",
       vertex.cex = 2,
       vertex.sides=(att$LEVEL+2)
 )

```{r Statnet: , echo=TRUE}

## Visualization: ggplot2
```{r Visualization: ggplot2}

# Dependencies
 library(network)
 library(sna)
 library(ggplot2)

library(visNetwork)
library(threejs)
 library(networkD3)
 library(ndtv)
 # ggnetwork - This package allows to pass network objects to ggplot2 and provides geometries to plot their elements.

 library(ggnetwork)

  # visualize your network graphs 
 library(GGally)

## [network visualization with ggplot2]

#
```{r Statnet: , echo=TRUE}

```{r igraph: , echo=TRUE}
